---
title: "R homework 1"
format: 
  html: 
   toc: true
   theme: 
     - united
   self-contained: true
   filters: 
     - honeypot
editor: visual
---

Use R to answer the questions below. For this assignment, you can write your answers to questions in comments in your R script, or you can submit written answers and R code in separate documents.

# Setup

Use the code below to load the tidyverse and import the NHANES data and the codebook with variable descriptions.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
nhanes<-read_csv("https://raw.githubusercontent.com/Neilblund/SDSB340-Spring-2026/refs/heads/master/Data/nhanes.csv")

codebook<-read_csv("https://raw.githubusercontent.com/Neilblund/SDSB340-Spring-2026/refs/heads/master/Data/nhanes_codebook.csv")



```

You can view the `codebook` to get additional detail on the variables in this data set.

## Question 1

What proportion of respondents **between ages 18 - 59** were regular marijuana users? (use `RegularMarij`)?

::: {.callout-note appearance="minimal"}
### Answer

The answer here could vary slightly depending on how you calculate the answer.

Here's an example of a solution that uses `filter` to eliminate people outside the age range, and then calculates the mean of `RegularMarij == "Yes"`. This approach would ignore anyone who didn't answer the question, so the proportion is \~28%

```{r}
nhanes|>
  filter(Age>=18 & Age<=59)|>
  summarise(mj = mean(RegularMarij == "Yes", na.rm=T))

```

You could also do this using `groupby`, but this wouldn't automatically drop `NA` responses, so the estimate is slightly lower at \~24%

```{r}
nhanes|>
  filter(Age>=18 & Age<=59)|>
  group_by(RegularMarij )|>
  summarize(n = n())|>
  ungroup()|>
  mutate(proportion = n/sum(n))

```
:::

## Question 2

Stage 2 hypertension is described as a Systolic blood pressure reading greater than or equal to 140, OR a diastolic reading greater than or equal to 90.

Using `BPSysAve` as your measure for systolic blood pressure and `BPDiaAve` as your measure for diastolic, use `mutate` to create a new variable called `stage2hyper` that is `TRUE` if a respondent has stage 2 hypertension and `FALSE` otherwise.

### Answer

::: {.callout-note appearance="minimal"}
You could do this with `mutate`

```{r}
nhanes<-nhanes|>
  mutate(stage2hyper = BPSysAve  >= 140 | BPDiaAve >=90)


```

We could also do this without `mutate`, the main difference is that we need to use the `$` notation here:

```{r}
nhanes$stage2hyper <- nhanes$BPSysAve  >= 140 | nhanes$BPDiaAve >=90

```
:::

## Question 3

Use `filter`, `groupby` , `summarize` , and `mean` to calculate the mean number of `DaysPhysHlthBad` for **adults** with and without stage 2 hypertension. Briefly discuss your results.

::: {.callout-note appearance="minimal"}
### Answer

The average for adults without stage 2 hypertension was 3.35, the average for adults with stage 2 hypertension was 4.14. Unsurprisingly, having very high blood pressure is associated with poor health.

```{r}
nhanes|>
  filter(Age >= 18)|>
  group_by(stage2hyper)|>
  summarize(DaysPhysHlthBad = mean(DaysPhysHlthBad, na.rm=T))

```
:::

## Question 4

Calculate the proportion of **adult respondents** who had blood pressure readings that indicated stage 2 hypertension by `Gender`.

Which group had a higher rate of hypertension?

::: {.callout-note appearance="minimal"}
### Answer

Stage 2 hypertension is about 2 percentage points higher among men compared to women (at least for those respondents who had a blood pressure reading).

```{r}
nhanes|>
  filter(Age>=18)|>
  group_by(Gender)|>
  summarize(stage2hyper= mean(stage2hyper, na.rm=T))
  

```

Fun fact: this relationship reverses after age 60 or so:

```{r}
nhanes|>
  filter(Age>=65)|>
  group_by(Gender)|>
  summarize(stage2hyper= mean(stage2hyper, na.rm=T))
  

```
:::

## Question 5

Use `filter`, `fct_relevel`, `group_by` and `summarize` to get the proportion of adult respondents who use a computer more than 4 hours per day (`CompHrsDay`) by `Education`. Ensure that values of education are displayed in a sensible order in your output.

Are more educated people more likely or less likely to spend more than 4 hours in front of the computer?

::: {.callout-note appearance="minimal"}
### Answer

As the education level increases, so do the changes of spending more than 4 hours in front of the TV.

```{r}

nhanes|>
  filter(Age>=18)|>
  mutate(Education = fct_relevel(Education, '8th Grade', 
                                 '9 - 11th Grade',
                                 'High School',
                                 'Some College',
                                 'College Grad'
                                 ))|>
  group_by(Education)|>
  summarize(CompHrsDay = mean(CompHrsDay%in%c('4_hr', 'More_4_hr'), 
                              na.rm=T))|>
  drop_na()


```
:::
